### 2023-04-02 outputs/paper.qmd
Initial commit

---
title: "Fast Fashion's Production Rate is Observedly Unsustainable: Why and How did we get here?"
author: "Jayden Jung"
thanks: "Code and data are available at: https://github.com/jj-andj/fast-fashion-analysis"
date: "April 21, 2023"
date-format: long
format: 
  pdf: 
    documentclass: article
    geometry: margin = 1in
    abstract: "An abstract is included and appropriately pitched to a non-specialist audience. The abstract answers: 1) what was done, 2) what was found, and 3) why this matters (all at a high level). Likely four sentences. Abstract must make clear what we learn about the world because of this paper."
output: 
  bookdown::pdf_document2
toc: false
fig_caption: yes
nocite: '@*'
bibliography: references.bib
---

# Introduction

> "Fast fashion isn't free. Someone, somewhere is paying."
>
> --- Lucy Siegle, Journalist and Producer of *The True Cost*, 2015.

1\) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found.

Estimand

# Data

Lorem

# Results

Lorem

# Discussion

## Findings

Lorem

## Why did we get here? Exploring culture-based roots of Fast Fashion

Lorem

## At what cost? Reviewing the ramifications of Fast Fashion

Lorem

## Defining Solution Spaces

Lorem

## Future Research

Lorem

## Limitations

Lorem

\clearpage

# References {.unnumbered}